CODECHECK: An Open Science initiative for the independent execution of computations underlying research articles during peer review to improve reproducibility

### 2021-02-23 codecheck-paper.tex
I'm done for now...

Codecheckers could be drawn from a regular pool of volunteer
\textbf{scientific reviewers}, or from a special group of
\textbf{reproducibility reviewers} via specific roles such as
\textbf{reproducibility editors}, or \textbf{editorial staff} with a
publisher.  A \emph{single codechecker} is sufficient to verify the
workflow since it is mostly a factual process and does not require much
debate or opinion.  Code usually harbours systematic and repeatable
mistakes and is thereby more reliable and auditable than processes
controlled by humans \cite{tibav:42484}, e.g.,~in a laboratory.  If
however publication of the paper depends on the CODECHECK, a second
opinion may be required.

@@

\textbf{Do workflows need to be codechecked multiple times?}  If a
paper is checked at the start of peer review, it might need
re-checking if the paper is modified during peer review.  This is
inevitable, and happened to us \cite{cert-2020-012}. This can
also be desired though, if interactions between author, reviewer,
and codechecker led to improvements.  Checking the manuscript the
second time is likely to be much less work than the first time.

@@

CODECHECK works --- we have created a considerable number of
certificates to demonstrate it.  The creation of certificates and
interactions with authors and editors shaped the principles and the
workflow and also confirmed the approach taken. This result
corroborates findings from similar evaluations of reproducible
computational research in journals and conferences.  CODECHECKs
increase transparency of the checked papers and can contribute to
building trust in scientific results.
% "science earns trust through transparency" > https://twitter.com/briannosek/status/1290240244101521408
The set of shared principles and common name, through recognition
value, will allow researchers to judge the level of scrutiny that
results have faced. CODECHECK requires direct acknowledgement of the
codechecker's contributions, not indirectly via citations of
reproductions or informal credit.

@@

While developed for the current ``paper''-centric publication process,
the CODECHECK principles would also work well with novel publication
paradigms, e.g., peer-reviewed computational notebooks
\cite{earthcube_new_2020}, iterative and granular communication of
scientific outputs, articles with live-code
\cite{perkel_pioneering_2019-1} such as eLife's ERA
(see \url{https://elifesciences.org/labs/dc5acbde/welcome-to-a-new-era-of-reproducible-publishing}),
decentralized infrastructure and public reviewer reputation systems
\cite{tenorio-fornes_towards_2019}, and completely new visions for
scholarly communication and peer review such as described by Amy J. Ko
in \emph{A modern vision for peer review}
(\url{https://medium.com/bits-and-behavior/a-modern-vision-for-peer-review-d5f73f0fae07}).
An explicit segmentation of research steps could even make the focus
of a CODECHECK easier by only checking the ``analysis''
sub-publication.  The discovery of CODECHECKs could be increased by
depositing certificates into public databases of reproductions, such as
\emph{SciGen.Report}.  Public researcher profiles, such as ORCID, may
consider different types of reviewer activity to capture how
independent code execution contributes to science.  Notably, the
discussed limitations are largely self-imposed for easier acceptance
and evolutionary integration, as to not break the current system and
increase demands gradually without leaving practitioners behind.

>>>>

Codecheckers could be drawn from a regular pool of
\textbf{peer-reviewers}, or from a special group of
\textbf{reproducibility reviewers} via specific roles such as
\textbf{reproducibility editors}, or \textbf{editorial staff} with a
publisher.  One codechecker is sufficient to verify the workflow since
it is mostly a factual process.  Code usually harbours systematic and
repeatable mistakes and is thereby more reliable and auditable than
processes controlled by humans \cite{tibav:42484}, e.g.,~in a
laboratory.  If however publication of the paper depends on the
CODECHECK, a second opinion may be required.

@@

\textbf{Do workflows need to be codechecked multiple times?}  If a
paper is checked at the start of peer review, it might need
re-checking if the paper is modified during peer review.  This is
inevitable, and happened to us \cite{cert-2020-012}. This is desirable
though, if interactions between author, reviewer, and codechecker led
to improvements.  Checking the manuscript the second time is likely to
be much less work than the first time.

@@

CODECHECK works --- we have created a considerable number of
certificates to demonstrate it.  The creation of certificates and
interactions with authors and editors shaped the principles and the
workflow and also confirmed the approach taken. This result
corroborates findings from similar evaluations of reproducible
computational research in journals and conferences.  CODECHECKs
increase transparency of the checked papers and can contribute to
building trust in research findings.
% "science earns trust through transparency" > https://twitter.com/briannosek/status/1290240244101521408
The set of shared principles and common name, through recognition
value, will allow researchers to judge the level of scrutiny that
results have faced. CODECHECK requires direct acknowledgement of the
codechecker's contributions, not indirectly via citations of
reproductions or informal credit.

@@

While developed for the current ``paper''-centric publication process,
the CODECHECK principles would also work well with novel publication
paradigms, e.g., peer-reviewed computational notebooks
\cite{earthcube_new_2020}, iterative and granular communication of
research outputs, articles with live-code
\cite{perkel_pioneering_2019-1} such as eLife's ERA
(see \url{https://elifesciences.org/labs/dc5acbde/welcome-to-a-new-era-of-reproducible-publishing}),
decentralized infrastructure and public reviewer reputation systems
\cite{tenorio-fornes_towards_2019}, and completely new visions for
scholarly communication and peer review such as described by Amy J. Ko
in \emph{A modern vision for peer review}
(\url{https://medium.com/bits-and-behavior/a-modern-vision-for-peer-review-d5f73f0fae07}).
An explicit segmentation of research steps could even make the focus
of a CODECHECK easier by only checking the ``analysis''
sub-publication.  The discovery of CODECHECKs could be increased by
depositing certificates into public databases of reproductions, such as
\emph{SciGen.Report}.  Public researcher profiles, such as ORCID, may
consider different types of reviewer activity to capture how
independent code execution contributes to science.  Notably, the
discussed limitations are largely self-imposed for easier acceptance
and evolutionary integration, as to not break the current system and
increase demands gradually without leaving practitioners behind.
